### 2023-04-05 main.tex
ii

At each site $x\in V$, we describe the occupation with an $N-$dimensional vector $n^{x}=(n_{1}^{x},\ldots,n_{N}^{x})$ in which the value of the $k$-th component $n_{a}^{x}$ denotes the number of particles of species $k\in \{1,\ldots,N\}$ at site $x\in V$. On each vertex we allow a total  maximal occupancy $\nu\in \mathbb{N}$. Then, the configuration space of the process on the graph $G$ is 

@@

when for any $k\in\{1,\ldots,N\}$

@@

$(a^{\dagger})_{a}^{\widehat{x}}$ denotes a copy of $a^{\dagger}$ acting on the extra-site $\widehat{x}$ and on the species $k\in\{2,\ldots,N\}$. \\

@@

\begin{equation}\label{transpositionPropertyH}
    H_{x,y}^{T}=\left(R_{x}R_{y}\right)H_{x,y}\left(R_{x}R_{y}\right)^{-1}
\end{equation}
therefore we have found the symilarity tranfsormation between $H_{x,y}$ and its transposed. \\
As already pointed out, $H_{x,y}$ is a linear function of $\Delta(C_{2})$. This implies that symmetries of  $\Delta(C_{2})$ are also symmetries of  $H_{x,y}$. Moreover, the coproduct is a Lie algebra homomorphism, i.e. $\forall A,B\in gl(N)$
\begin{equation}\label{symmetryCoproduct}
    [A,B]=0\quad \Longrightarrow\quad \left[\Delta(A),\Delta(B)\right]=0
\end{equation}
This tells us that it is enough to find a symmetry of $C_{2}$. This element belongs to the center of $gl(N)$, and therefore it commutes with all the element of this algebra. Consider
\begin{equation}\label{equationE}
    E=\sum_{a=2}^{N}E_{a1}
\end{equation}
To obtain a product structure of the elements of the duality matrix, we introduce
\begin{equation}
    S=\exp{(\Delta(E))}
\end{equation}
Finally, for fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
    S_{x,y}=\exp{(\Delta^{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta^{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. This operator $S^{x,y}$ satisfies
\begin{equation}\label{symmetryH}
    \left[S_{x,y},H_{x,y}\right]=0
\end{equation}
i.e. $S_{x,y}$ is a symmetry of $H_{x,y}$. \\ Exploiting these considerations we write
\begin{equation}
    \begin{split}
        H_{x,y}^{T}D&=(R_{x}R_{y})H_{x,y}(R_{x}R_{y})^{-1}\left(d_{x}\otimes\dd\right)\left(d_{y}\otimes\mathcal{D}_{\widehat{y}}\right)\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\widehat{z}}\right)
        \\&=\left(R_{x}\exp{(E^{x})}\otimes \dd\right)\left(R_{y}\exp{(E^{y})}\otimes \mathcal{D}_{\widehat{y}}\right)H_{x,y}\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\widehat{z}}\right)
        \\&=
        DH_{x,y}
    \end{split}
\end{equation}
 where we used \eqref{transpositionPropertyH} and \eqref{symmetryH} in the second equality. Thus, \eqref{edgeDualRealtion} is proved. 
 %\textbf{Reamrk}: it is important to notice that the elements of the duality matrix \eqref{dualityElements} are well defined only if at each site $x\in V$ and for every species $k\in \{2,\ldots,N\}$ the number of dual particles $r_{x}^{a}$ is lower or equal than the number of original particles $n_{a}^{x}$. This implies that the dynamics of the dual process is simpler, because described by a lower number of particles. By the way, it is possible to perform computations similar to the one made in this proof choosing instead of  \eqref{equationE} {\color{red} $E=\sum_{a=2}^{N}E_{1a}$ or $E=\sum_{a=2}^{N}E_{aa}$.}
 %{\color{blue} That's confusing. What is E?}

@@

Recalling the action of the bosonic creation operator acting at site $\widehat{x}$ and on the species $k\in \{2,\ldots,N\}$ we have that 

@@

Using \eqref{bosonicKX} by a fixed $k\in \{2,\ldots,N\}$ we rewrite the term of \eqref{siteHadamardII} with Lie generator $E_{1a}^{x}$ on the right hand side of \eqref{siteHadamardII} as 

@@

	\frac{\Gamma(2+\sum_{i=1}^{L}n_{1}^{x})}{\Gamma(2+L)}=\frac{\Gamma\left(2+\sum_{i=1}^{L}\delta_{s_{x},1}\right)}{\Gamma\left(L+2\right)}

@@

        S_{2}^{-1}=\sum_{s_{1}=1}^{N}e_{i}^{x}+Q

@@

	(\mathcal{S}_{2}^{-1})_{\bm{s}}=\left(S_{2}^{-1}\right)^{\otimes L}_{\bm{s}}&=\,e_{s_{1},\ldots,s_{L}}^{s_{1},\ldots,s_{L}}+\sum_{q_{1}=1}^{N}(1-\delta_{s_{q_{1}},1})e_{s_{1},\ldots,s_{q_{1}},\ldots,s_{L}}^{s_{1},\ldots,1,\ldots,s_{L}}\beta_{s_{q_{1}}}\\&+\sum_{q_{1},q_{2}=1\,:\,q_{1}\neq q_{2}}^{L}(1-\delta_{s_{q_{1}},1})(1-\delta_{s_{q_{2}},1})e_{s_{1},\ldots,s_{q_{1}},\ldots,s_{q_{2}}\ldots,s_{L}}^{s_{1},\ldots,1,\ldots,1,\ldots,s_{L}}\beta_{s_{q_{1}}}\beta_{s_{q_{2}}}

@@

where $D(\bm{n},\bm{\xi})$ is the duality function defined in \eqref{dualityFunctionChain}. Here, we denote by $|\xi_{a}|=\sum_{x=1}^{L}\xi_{a}^{x}$ and by $i_{a}$ the number of dual particles of species $k\in\{2,\ldots,N\}$ absorbed at left, thus we denote by $|\xi_{a}|-i_{a}$ the number of particles of the same type absorbed at right.\\

@@

where with the symbol $\Delta_{N}$ we denote the discrete Laplacian operator on the chain, i.e. $\Delta_{L}n_{a}^{x}=n_{a}^{x+1}+n_{a}^{x-1}-2n_{a}^{x}$, $\forall x\in \{2,\ldots,L-1\}$ and $\forall k\in \{2,\ldots,N\}$. In the spirit of \cite{casini2022uphill} we can write he the Master's equation for the evolution of the average density occupation, obtaining a system of $N-1$ difference differential equation where the discrete Laplacian of all species are coupled and linear reaction term are present. \\ 
The process defined by generator \eqref{RDGenerator} has reversible product measure when all the reservoirs have the same parameter for each site and each species, i.e. $\alpha_{a}^{x}=\alpha$, $\forall k\in \{1,\ldots,N\}$ and $x\in V$. This measure is given by

>>>>

At each site $x\in V$, we describe the occupation with an $N-$dimensional vector $n^{x}=(n_{1}^{x},\ldots,n_{N}^{x})$ in which the value of the $k$-th component $n_{a}^{x}$ denotes the number of particles of species $a\in \{1,\ldots,N\}$ at site $x\in V$. On each vertex we allow a total  maximal occupancy $\nu\in \mathbb{N}$. Then, the configuration space of the process on the graph $G$ is 

@@

when for any $a\in\{1,\ldots,N\}$

@@

$(a^{\dagger})_{a}^{\widehat{x}}$ denotes a copy of $a^{\dagger}$ acting on the extra-site $\widehat{x}$ and on the species $a\in\{2,\ldots,N\}$. \\

@@

\begin{equation}\label{transpositionPropertyH}
    H_{x,y}^{T}=\left(R_{x}R_{y}\right)H_{x,y}\left(R_{x}R_{y}\right)^{-1}
\end{equation}
therefore we have found the symilarity tranfsormation between $H_{x,y}$ and its transposed. \\
As already pointed out, $H_{x,y}$ is a linear function of $\Delta(C_{2})$. This implies that symmetries of  $\Delta(C_{2})$ are also symmetries of  $H_{x,y}$. Moreover, the coproduct is a Lie algebra homomorphism, i.e. $\forall A,B\in gl(N)$
\begin{equation}\label{symmetryCoproduct}
    [A,B]=0\quad \Longrightarrow\quad \left[\Delta(A),\Delta(B)\right]=0
\end{equation}
This tells us that it is enough to find a symmetry of $C_{2}$. This element belongs to the center of $gl(N)$, and therefore it commutes with all the element of this algebra. Consider
\begin{equation}\label{equationE}
    E=\sum_{a=2}^{N}E_{a1}
\end{equation}
To obtain a product structure of the elements of the duality matrix, we introduce
\begin{equation}
    S=\exp{(\Delta(E))}
\end{equation}
Finally, for fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
    S_{x,y}=\exp{(\Delta^{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta^{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. This operator $S^{x,y}$ satisfies
\begin{equation}\label{symmetryH}
    \left[S_{x,y},H_{x,y}\right]=0
\end{equation}
i.e. $S_{x,y}$ is a symmetry of $H_{x,y}$. \\ Exploiting these considerations we write
\begin{equation}
    \begin{split}
        H_{x,y}^{T}D&=(R_{x}R_{y})H_{x,y}(R_{x}R_{y})^{-1}\left(d_{x}\otimes\dd\right)\left(d_{y}\otimes\mathcal{D}_{\widehat{y}}\right)\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\widehat{z}}\right)
        \\&=\left(R_{x}\exp{(E^{x})}\otimes \dd\right)\left(R_{y}\exp{(E^{y})}\otimes \mathcal{D}_{\widehat{y}}\right)H_{x,y}\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{\widehat{z}}\right)
        \\&=
        DH_{x,y}
    \end{split}
\end{equation}
 where we used \eqref{transpositionPropertyH} and \eqref{symmetryH} in the second equality. Thus, \eqref{edgeDualRealtion} is proved. 
 %\textbf{Reamrk}: it is important to notice that the elements of the duality matrix \eqref{dualityElements} are well defined only if at each site $x\in V$ and for every species $a\in \{2,\ldots,N\}$ the number of dual particles $r_{x}^{a}$ is lower or equal than the number of original particles $n_{a}^{x}$. This implies that the dynamics of the dual process is simpler, because described by a lower number of particles. By the way, it is possible to perform computations similar to the one made in this proof choosing instead of  \eqref{equationE} {\color{red} $E=\sum_{a=2}^{N}E_{1a}$ or $E=\sum_{a=2}^{N}E_{aa}$.}
 %{\color{blue} That's confusing. What is E?}

@@

Recalling the action of the bosonic creation operator acting at site $\widehat{x}$ and on the species $a\in \{2,\ldots,N\}$ we have that 

@@

Using \eqref{bosonicKX} by a fixed $a\in \{2,\ldots,N\}$ we rewrite the term of \eqref{siteHadamardII} with Lie generator $E_{1a}^{x}$ on the right hand side of \eqref{siteHadamardII} as 

@@

	\frac{\Gamma(2+\sum_{x=1}^{L}n_{1}^{x})}{\Gamma(2+L)}=\frac{\Gamma\left(2+\sum_{x=1}^{L}\delta_{s_{x},1}\right)}{\Gamma\left(L+2\right)}

@@

        S_{2}^{-1}=\sum_{s_{1}=1}^{N}e_{s_{1}}^{s_{1}}+Q

@@

	(\mathcal{S}_{2}^{-1})_{\bm{s}}=\left(S_{2}^{-1}\right)^{\otimes L}_{\bm{s}}&=\,e_{s_{1},\ldots,s_{L}}^{s_{1},\ldots,s_{L}}+\sum_{q_{1}=1}^{L}(1-\delta_{s_{q_{1}},1})e_{s_{1},\ldots,s_{q_{1}},\ldots,s_{L}}^{s_{1},\ldots,1,\ldots,s_{L}}\beta_{s_{q_{1}}}\\&+\sum_{q_{1},q_{2}=1\,:\,q_{1}\neq q_{2}}^{L}(1-\delta_{s_{q_{1}},1})(1-\delta_{s_{q_{2}},1})e_{s_{1},\ldots,s_{q_{1}},\ldots,s_{q_{2}}\ldots,s_{L}}^{s_{1},\ldots,1,\ldots,1,\ldots,s_{L}}\beta_{s_{q_{1}}}\beta_{s_{q_{2}}}

@@

where $D(\bm{n},\bm{\xi})$ is the duality function defined in \eqref{dualityFunctionChain}. Here, we denote by $|\xi_{a}|=\sum_{x=1}^{L}\xi_{a}^{x}$ and by $i_{a}$ the number of dual particles of species $a\in\{2,\ldots,N\}$ absorbed at left, thus we denote by $|\xi_{a}|-i_{a}$ the number of particles of the same type absorbed at right.\\

@@

where with the symbol $\Delta_{N}$ we denote the discrete Laplacian operator on the chain, i.e. $\Delta_{L}n_{a}^{x}=n_{a}^{x+1}+n_{a}^{x-1}-2n_{a}^{x}$, $\forall x\in \{2,\ldots,L-1\}$ and $\forall a\in \{2,\ldots,N\}$. In the spirit of \cite{casini2022uphill} we can write he the Master's equation for the evolution of the average density occupation, obtaining a system of $N-1$ difference differential equation where the discrete Laplacian of all species are coupled and linear reaction term are present. \\ 
The process defined by generator \eqref{RDGenerator} has reversible product measure when all the reservoirs have the same parameter for each site and each species, i.e. $\alpha_{a}^{x}=\alpha$, $\forall a\in \{1,\ldots,N\}$ and $x\in V$. This measure is given by
